Combinatorial Mixture Models for Single-Cell Assays with Application to Vaccine Studies

### 2012-07-30 MIMOSA.tex
Resized Figure 1. Enlarged subpanel labels (A,B,C etc)
Signed-off-by: Greg Finak <gfinak@fhcrc.org>

\begin{figure} %  figure placement: here, top, bottom, or page
   \centering
\begin{tikzpicture} [auto, node distance=0cm]
\node at (0,0) (A){
    \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] at (0,0) (foo) {\includegraphics[width=.6\columnwidth]{Figures/2}};
    \begin{scope}[x={(foo.south east)},y={(foo.north west)}]
    \node at (0,1) [font=\tiny\sffamily] {A} ;
    \node at (0.5,1) [font=\tiny\sffamily] {B} ;
    \end{scope}
\end{tikzpicture}
 };
 \node [below=of A] (B) {
 \begin{tikzpicture}
    \node[anchor=south west, inner sep=0] at (0,0) (bar){\includegraphics[width=.6\columnwidth]{Figures/1}};
    \begin{scope} [x={(bar.south east)},y={(bar.north west)}]
    \node at (0,1) [font=\tiny\sffamily] {C} ;
    \node at (0.5,1) [font=\tiny\sffamily] {D} ;
    \end{scope}
    \end{tikzpicture}
 };
\end{tikzpicture}
   \caption{Performance of MIMOSA (EM and MCMC implementations, one-sided model) and competing methods on ICS data from HVTN065. Sensitivity and specificity (ROC analysis) as well as observed and nominal false discovery rates for positivity calls from CD4+ T-cells stimulated with A-B) ENV-1-PTEG and expressing IFNg or C-D) ENV-1-PTEG and expressing IL2. ROC and FDR plots of other cytokine combinations can be found in Web Figure A.}
\label{fig:HVTN065}
\end{figure}

@@

\begin{figure}
\centering
\begin{tikzpicture} [auto,node distance=0cm]
\node at (0,0) (A) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_plot1.pdf}};
\begin{scope} [x={(foo.south east)},y={(foo.north west)}]
\node at (0,1) [font=\tiny\sffamily] {A} ;
\end{scope}
\end{tikzpicture}
};
\node [right=of A] (B) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_plot2.pdf}};
\begin{scope} [x={(bar.south east)},y={(bar.north west)}]
\node at (-0.1,1.1) [font=\tiny\sffamily] {B} ;
\end{scope}
\end{tikzpicture}
};
\node [below=of A] (C) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (baz) {\includegraphics[width=0.4\columnwidth]{Figures/FluidigmFisher.pdf}};
\begin{scope} [x={(baz.south east)},y={(baz.north west)}]
\node at (0,1) [font=\tiny\sffamily] {C} ;
\end{scope}
\end{tikzpicture}
};
\end{tikzpicture}
\caption{Signed posterior probability, difference and log-odds ratio of the proportion of single-cells expressing each gene on a 96x96 Fluidigm array. The posterior probability of response times the sign of the change in expression is shown in A) (red indicates a decrease, green an increase, relative to the control). Columns and rows are clustered based on these signed posterior probabilities. B) The posterior differences in proportion of cells expressing a gene in the stimulated vs. control samples. Rows and columns are ordered as in A) for comparison. The traces show the deviations of each cell from zero. Colors along the columns denote different stimulations (green: CMV pp65 nlv5, red: HIV Gag, orange: HIV Nef, yellow: CMV pp65 tm10). C) Clustering of the signed q-values from Fisher's exact test. Genes selected from Fisher's exact test at the 10\% FDR level. }
\label{fig:fluidigm}
\end{figure}

@@

\begin{figure} %  figure placement: here, top, bottom, or page
   \centering
   \begin{tikzpicture} [auto, node distance=0cm]
   \node at (0,0) (A) {
\begin{tikzpicture}
    \node [anchor=south west] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_onesided_ROC_5000.pdf}};
    \begin{scope}[x={(foo.south east)},y={(foo.north west)}]
        \node at (0,1) [font=\tiny\sffamily] {A} ;
        \end{scope}
    \end{tikzpicture}
    };
    \node [right=of A] (B) {
    \begin{tikzpicture}
    \node [anchor=south west] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_onesided_FDR_5000.pdf}};
    \begin{scope}[x={(bar.south east)},y={(bar.north west)}]
    \node at (0,1) [font=\tiny\sffamily] {B} ;
    \end{scope}
    \end{tikzpicture}
    };
        \node [below=of A] (C) {
    \begin{tikzpicture}
    \node [anchor=south west] at (0,0) (c) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_twosided_ROC_5000.pdf}};
    \begin{scope}[x={(c.south east)},y={(c.north west)}]
    \node at (0,1) [font=\tiny\sffamily] {C} ;
    \end{scope}
    \end{tikzpicture}
    };
    \node [right=of C] (D) {
    \begin{tikzpicture}
    \node [anchor=south west] at (0,0) (d) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_onesided_FDR_5000.pdf}};
    \begin{scope}[x={(d.south east)},y={(d.north west)}]
    \node at (0,1) [font=\tiny\sffamily] {D} ;
    \end{scope}
    \end{tikzpicture}
    };

@@

\begin{figure}
\centering
\begin{tikzpicture} [auto, node distance=0cm]
\node at (0,0) (A) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_BIRC3_CCL5_Unstim_Counts.pdf}};
\begin{scope} [x={(foo.south east)},y={(foo.north west)}]
\node at (0,1) [font=\tiny\sffamily] {A} ;
\end{scope}
\end{tikzpicture}
};
\node [right=of A] (B){
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_BIRC3_CCL5_Stim_Counts.pdf}};
\begin{scope} [x={(bar.south east)},y={(bar.north west)}]
\node at (0,1) [font=\tiny\sffamily] {B} ;
\end{scope}
\end{tikzpicture}
};
%\node [anchor = south west, inner sep=0] at (0,-4) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_Marginalized_BIRC3_CCL5.pdf}};
%\node [anchor=south west, inner sep=0] at (6.5,-4) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_BIRC3_CCL5.pdf}};
%\node at (0,0) [font=\tiny\sffamily] {C} ;
%\node at (6.25,0) [font=\tiny\sffamily] {D} ;
\end{tikzpicture}
\caption{Counts of cells expressing different combinations of BIRC3 and CCL5 genes in the A) unstimulated and B) stimulated conditions. No difference is observed from the marginalized counts, while multivariate MIMOSA detects a difference between stimulated and unstimulated conditions in 13 of 16 samples. Sample names highlighted in red identify those where MIMOSA did not detect a difference.}
\label{fig:polyfunctionality}
\end{figure}

@@

\begin{figure} %  figure placement: here, top, bottom, or page
   \centering
\begin{tikzpicture} [auto, node distance=0cm]
\node at (0,0) (A) {
\begin{tikzpicture}
    \node[anchor=south west, inner sep=0] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Multivariate_Sims_ROC_rev.pdf}};
\begin{scope} [x={(foo.south east)},y={(foo.north west)}]
    \node at (0,1) [font=\tiny\sffamily] {A} ;
\end{scope}
\end{tikzpicture}
};
\node [right=of A] (B) {
\begin{tikzpicture}
   \node[anchor=south west, inner sep=0] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Multivariate_Sims_FDR_rev.pdf}};
   \begin{scope} [x={(bar.south east)},y={(bar.north west)}]
       \node at (0,1) [font=\tiny\sffamily] {B} ;
\end{scope}
\end{tikzpicture}
};
\end{tikzpicture}
   \caption{Multivariate simulations from a two-sided model. Ten, eight-dimensional data sets were simulated from a two-sided model with an effect sizes of $2.5\times 10^{-3}$ and $-2.5\times 10^{-3}$ in two of the eight dimensions (N=1,500). Multivariate MIMOSA was compared against Fisher's exact test, and the likelihood ratio test. A) Average ROC curves for the competing methods over 10 simulations. B) Average observed and nominal false discovery rate for each method over 10 simulations.}
   \label{fig:mvsimulations}
\end{figure}

>>>>

\begin{figure} %  figure placement: here, top, bottom, or page
   \centering
\begin{tikzpicture} [auto, node distance=0cm]
\node at (0,0) (A){
    \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] at (0,0) (foo) {\includegraphics[width=\columnwidth]{Figures/2}};
    \begin{scope}[x={(foo.south east)},y={(foo.north west)}]
    \node at (0,1) [font=\small\sffamily] {A} ;
    \node at (0.5,1) [font=\small\sffamily] {B} ;
    \end{scope}
\end{tikzpicture}
 };
 \node [below=of A] (B) {
 \begin{tikzpicture}
    \node[anchor=south west, inner sep=0] at (0,0) (bar){\includegraphics[width=\columnwidth]{Figures/1}};
    \begin{scope} [x={(bar.south east)},y={(bar.north west)}]
    \node at (0,1) [font=\small\sffamily] {C} ;
    \node at (0.5,1) [font=\small\sffamily] {D} ;
    \end{scope}
    \end{tikzpicture}
 };
\end{tikzpicture}
   \caption{Performance of MIMOSA (EM and MCMC implementations, one-sided model) and competing methods on ICS data from HVTN065. Sensitivity and specificity (ROC analysis) as well as observed and nominal false discovery rates for positivity calls from CD4+ T-cells stimulated with A-B) ENV-1-PTEG and expressing IFNg or C-D) ENV-1-PTEG and expressing IL2. ROC and FDR plots of other cytokine combinations can be found in Web Figure A.}
\label{fig:HVTN065}
\end{figure}

@@

\begin{figure}
\centering
\begin{tikzpicture} [auto,node distance=0cm]
\node at (0,0) (A) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_plot1.pdf}};
\begin{scope} [x={(foo.south east)},y={(foo.north west)}]
\node at (0,1) [font=\small\sffamily] {A} ;
\end{scope}
\end{tikzpicture}
};
\node [right=of A] (B) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_plot2.pdf}};
\begin{scope} [x={(bar.south east)},y={(bar.north west)}]
\node at (-0.1,1.1) [font=\small\sffamily] {B} ;
\end{scope}
\end{tikzpicture}
};
\node [below=of A] (C) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (baz) {\includegraphics[width=0.4\columnwidth]{Figures/FluidigmFisher.pdf}};
\begin{scope} [x={(baz.south east)},y={(baz.north west)}]
\node at (0,1) [font=\small\sffamily] {C} ;
\end{scope}
\end{tikzpicture}
};
\end{tikzpicture}
\caption{Signed posterior probability, difference and log-odds ratio of the proportion of single-cells expressing each gene on a 96x96 Fluidigm array. The posterior probability of response times the sign of the change in expression is shown in A) (red indicates a decrease, green an increase, relative to the control). Columns and rows are clustered based on these signed posterior probabilities. B) The posterior differences in proportion of cells expressing a gene in the stimulated vs. control samples. Rows and columns are ordered as in A) for comparison. The traces show the deviations of each cell from zero. Colors along the columns denote different stimulations (green: CMV pp65 nlv5, red: HIV Gag, orange: HIV Nef, yellow: CMV pp65 tm10). C) Clustering of the signed q-values from Fisher's exact test. Genes selected from Fisher's exact test at the 10\% FDR level. }
\label{fig:fluidigm}
\end{figure}

@@

\begin{figure} %  figure placement: here, top, bottom, or page
   \centering
   \begin{tikzpicture} [auto, node distance=0cm]
   \node at (0,0) (A) {
\begin{tikzpicture}
    \node [anchor=south west] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_onesided_ROC_5000.pdf}};
    \begin{scope}[x={(foo.south east)},y={(foo.north west)}]
        \node at (0,1) [font=\small\sffamily] {A} ;
        \end{scope}
    \end{tikzpicture}
    };
    \node [right=of A] (B) {
    \begin{tikzpicture}
    \node [anchor=south west] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_onesided_FDR_5000.pdf}};
    \begin{scope}[x={(bar.south east)},y={(bar.north west)}]
    \node at (0,1) [font=\small\sffamily] {B} ;
    \end{scope}
    \end{tikzpicture}
    };
        \node [below=of A] (C) {
    \begin{tikzpicture}
    \node [anchor=south west] at (0,0) (c) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_twosided_ROC_5000.pdf}};
    \begin{scope}[x={(c.south east)},y={(c.north west)}]
    \node at (0,1) [font=\small\sffamily] {C} ;
    \end{scope}
    \end{tikzpicture}
    };
    \node [right=of C] (D) {
    \begin{tikzpicture}
    \node [anchor=south west] at (0,0) (d) {\includegraphics[width=0.4\columnwidth]{Figures/Sim_onesided_FDR_5000.pdf}};
    \begin{scope}[x={(d.south east)},y={(d.north west)}]
    \node at (0,1) [font=\small\sffamily] {D} ;
    \end{scope}
    \end{tikzpicture}
    };

@@

\begin{figure}
\centering
\begin{tikzpicture} [auto, node distance=0cm]
\node at (0,0) (A) {
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_BIRC3_CCL5_Unstim_Counts.pdf}};
\begin{scope} [x={(foo.south east)},y={(foo.north west)}]
\node at (0,1) [font=\small\sffamily] {A} ;
\end{scope}
\end{tikzpicture}
};
\node [right=of A] (B){
\begin{tikzpicture}
\node[anchor=south west, inner sep=0] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_BIRC3_CCL5_Stim_Counts.pdf}};
\begin{scope} [x={(bar.south east)},y={(bar.north west)}]
\node at (0,1) [font=\small\sffamily] {B} ;
\end{scope}
\end{tikzpicture}
};
%\node [anchor = south west, inner sep=0] at (0,-4) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_Marginalized_BIRC3_CCL5.pdf}};
%\node [anchor=south west, inner sep=0] at (6.5,-4) {\includegraphics[width=0.4\columnwidth]{Figures/Fluidigm_Multivariate_BIRC3_CCL5.pdf}};
%\node at (0,0) [font=\small\sffamily] {C} ;
%\node at (6.25,0) [font=\small\sffamily] {D} ;
\end{tikzpicture}
\caption{Counts of cells expressing different combinations of BIRC3 and CCL5 genes in the A) unstimulated and B) stimulated conditions. No difference is observed from the marginalized counts, while multivariate MIMOSA detects a difference between stimulated and unstimulated conditions in 13 of 16 samples. Sample names highlighted in red identify those where MIMOSA did not detect a difference.}
\label{fig:polyfunctionality}
\end{figure}

@@

\begin{figure} %  figure placement: here, top, bottom, or page
   \centering
\begin{tikzpicture} [auto, node distance=0cm]
\node at (0,0) (A) {
\begin{tikzpicture}
    \node[anchor=south west, inner sep=0] at (0,0) (foo) {\includegraphics[width=0.4\columnwidth]{Figures/Multivariate_Sims_ROC_rev.pdf}};
\begin{scope} [x={(foo.south east)},y={(foo.north west)}]
    \node at (0,1) [font=\small\sffamily] {A} ;
\end{scope}
\end{tikzpicture}
};
\node [right=of A] (B) {
\begin{tikzpicture}
   \node[anchor=south west, inner sep=0] at (0,0) (bar) {\includegraphics[width=0.4\columnwidth]{Figures/Multivariate_Sims_FDR_rev.pdf}};
   \begin{scope} [x={(bar.south east)},y={(bar.north west)}]
       \node at (0,1) [font=\small\sffamily] {B} ;
\end{scope}
\end{tikzpicture}
};
\end{tikzpicture}
   \caption{Multivariate simulations from a two-sided model. Ten, eight-dimensional data sets were simulated from a two-sided model with an effect sizes of $2.5\times 10^{-3}$ and $-2.5\times 10^{-3}$ in two of the eight dimensions (N=1,500). Multivariate MIMOSA was compared against Fisher's exact test, and the likelihood ratio test. A) Average ROC curves for the competing methods over 10 simulations. B) Average observed and nominal false discovery rate for each method over 10 simulations.}
   \label{fig:mvsimulations}
\end{figure}
